---
title: "exercise-11"
format: html
editor: visual
---

## vload packages

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(MuMIn)
```

## load data

```{r}
d <- read_tsv("https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt", col_names = TRUE)
```

## step1

```{r}
d[d == -999.00] <- NA
```

## step2

```{r}
d <- d %>%
  select(-c(`litter size`, refs))
```

## step3

```{r}
d <- d %>%
  mutate(across(where(is.numeric), log))
```

## step4

```{r}
d <- d %>%
  mutate(relGest = residuals(lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)),
    relWean = residuals(lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)),
    relAFR = residuals(lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)),
    relLife = residuals(lm(`max. life(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)),
    relNewbornMass = residuals(lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)),
    relWeaningMass = residuals(lm(`wean mass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)))
```

## step5

```{r, warning=FALSE}
ggplot(d, aes(x = order, y = relLife)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Primates has the highest residual lifespan.

```{r, warning=FALSE}
ggplot(d, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Carnivora has the highest residual newborn mass.

```{r,warning=FALSE}
ggplot(d, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Perissodactyla has the highest residual weaning mass.

## step6
#### max. life(mo)
```{r}
predictors <- c("gestation(mo)", "newborn(g)", "weaning(mo)", "wean mass(g)", "litters/year", "mass(g)")

d_life <- d %>%
  drop_na(`max. life(mo)`, all_of(predictors))

options(na.action = na.fail)

m_life <- lm(`max. life(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `wean mass(g)` + `litters/year` + `mass(g)`, data = d_life)

(mods_life <- dredge(m_life))
```
Here, the best model is `max. life(mo)` ~ `gestation(mo)` + `weaning(mo)` + `litters/year` + `mass(g)`. Five models have a delta of 4 or less. Gestation(mo), litters/year and weaning(mo) appear in all top models.
```{r}
top_life <- get.models(mods_life, subset = delta < 4)
avg_life <- model.avg(top_life)
summary(avg_life)
confint(avg_life)
plot(avg_life, full = TRUE, intercept = FALSE)
```

#### AFR(mo)
```{r}
d_arf <- d %>%
  drop_na(`AFR(mo)`, all_of(predictors))

options(na.action = na.fail)

m_arf <- lm(`AFR(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `wean mass(g)` + `litters/year` + `mass(g)`, data = d_arf)

(mods_arf <- dredge(m_arf))
```
Here the best model is `AFR(mo)` ~ `gestation(mo)` +  `weaning(mo)` + `litters/year` + `mass(g)`. 7 models have a delta of 4 or less. Gestation(mo), litters/year and weaning(mo) appear in all top models.
```{r}
top_arf <- get.models(mods_arf, subset = delta < 4)
avg_arf <- model.avg(top_arf)
summary(avg_arf)
confint(avg_arf)
plot(avg_arf, full = TRUE, intercept = FALSE)
```